---
title: "Using audioBlast with R"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{audioblast}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE}
library(sonicscrewdriver)
```

## What is audioblast?

[Audioblast](https://audioblast.org) is a discovery tool for bioacoustic and ecoacoustic recordings. You can use it to search for recordings (e.g. by species, location, date), annotations of recordings, analysis results and traits.

Audioblast is currently hosted by the [Natural History Museum](https://www.nhm.ac.uk/) in London. It was conceived by [Ed Baker](https://ebaker.me.uk) at the [University of York](https://www.york.ac.uk/) as part of the Leverhulme Trust funded [Automated Acoustic Observatories](https://ebaker.me.uk/aao) project.

## Accessing audioblast

The `audioblast()` function of `SonicScrewdriveR` is a wrapper allowing full access to the [Audioblast API](https://api.audioblast.org). It is also possible to download many recordings through audioblast using `audioblastDownload()`.
