---
title: "Managing audio in R"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Managing audio in R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, results=FALSE}
library(sonicscrewdriver)
```

## Managing audio in R with SonicScrewdriveR

### Reading audio files
Several functions are available to read audio files into R, including the `readWave()` and `readMP3()` functions from the `tuneR` package, as well as tools from the package `av`. SonicScrewdriveR simplifies reading audio files by providing a single wrapper for these functions, `read_audio()`, which can read audio files in a variety of formats, including WAV, MP3, and FLAC. 

```{r}
filename <- system.file("extdata", "AUDIOMOTH.WAV", package="sonicscrewdriver")
w <- readAudio(filename)
```

### Performing analyses on channels

Some existing functions only operate on a single channel at a time. This may cause unnecessarily complex workflows when bulk analysing files with different numbers of channels. The `allChannels()` function applies a function to each channel and returns a list of analyses. This technique allows for the same analysis to be performed on each channel, without reference to the number of channels in the audio file. Optionally, a cluster can be specified to process channels on separate processor cores to increase analysis speed.

### Windowing

It is often desirable to process a long audio file in chunks. The `windowing()` function can be used to split an audio file into overlapping or non-overlapping windows.
