---
title: "Augmenting audio data in R with SonicScrewdriveR"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, results=FALSE}
library(sonicscrewdriver)
```

# Introduction

Augmenting data is a common practice in machine learning. It is the process of creating new data from existing data (e.g. by adding noise). This is done to increase the size of the training data and to make the model more robust. This article will show how to augment audio data in R using the `sonicscrewdriver` package.

All of the `generateX()` functions in `sonicscrewdriver` are designed to operate on `Wave`-like objects (`Wave` or `WaveMC` from `tuneR` of their Tagged equivalents) or a list of `Wave`-like objects. Similarly, all of these functions return a list of `Wave`-like objects. This means that you can chain these functions together to create complex data augmentation pipelines.

# Types of augmentation
In order to demonstrate the data augmentation process, we will use the `sonicscrewdriver` package to generate a Dirac pulse.

```{r}
# Generate a Dirac pulse
p <- pulse("dirac")

# Plot the pulse
seewave::oscillo(p)
```
## Noise augmentation

Noise augmentation is a common technique used to increase the size of the training data. It is the process of adding noise to the original data. The `sonicscrewdriver` package provides the `generateNoise()` function to add noise to audio data.

In this example we will add two different amounts of white noise to the Dirac pulse.

```{r, echo=FALSE}
augmented <- c(
  generateNoise(p, "white", noise.ratio=0.1),
  generateNoise(p, "white", noise.ratio=0.3)
)

for (i in 1:length(augmented)) {
  seewave::oscillo(augmented[[i]])
}

```

## Time shifting (rotation/padding)

Time shifting is the process of shifting the audio data by a certain number of samples. The `sonicscrewdriver` package provides the `generateTimeShift()` function to rotate audio data, either rotating the audio data within the file, or padding the file with silence.

To demonstrate time shifting we will generate the sum of a sine wave and a Dirac pulse.

```{r, echo=FALSE}
p1 <- tuneR::sine(440, duration=44100*3)
p2 <- pulse("dirac", duration=44100*3)

p <- 0.25*p1 + tuneR::normalize(p2)

seewave::oscillo(p)
```

Time shifting by rotation

```{r, echo=FALSE}
# Rotate the audio data by one second
rotated <- generateTimeShift(p, amount=1)

seewave::oscillo(rotated[[1]])
```

Alternatively, we can rotate the audio data.

```{r, echo=FALSE}
# Rotate the audio data by one second
rotated <- generateTimeShift(p, type="rotate", amount=1)

seewave::oscillo(rotated[[1]])
```

# Combining augmentations

The `sonicscrewdriver` package allows you to chain together multiple augmentations to create complex data augmentation pipelines. For example, you can add noise to the audio data and then rotate the audio data.

```{r, echo=FALSE}
# Load some test audio data
f <- system.file("extdata", "AUDIOMOTH.flac", package="sonicscrewdriver")
w <- readAudio(f)

# Add noise to the audio data and then rotate the audio data using pipes
augmented <- 
  w |> 
  generateNoise("white", noise.ratio=0.1) |> 
  generateTimeShift(amount=1)

# Create list of data and augmentations
data <- c(w, augmented)
```

It is possible to use anonymous functions to add multiple augmentations in a single step. This can be effectively used to generate large amounts of augmented data. The code below generates three different noise augmented versions of an input, and then two different time-shifted versions of each of those, yielding six augmented versions.

```{r, echo=FALSE}
# Load some test audio data
f <- system.file("extdata", "AUDIOMOTH.flac", package="sonicscrewdriver")
w <- readAudio(f)

# Add noise to the audio data and then rotate the audio data using pipes

augmented <- 
  w |>
  {\(x) c(generateNoise(x, "white", noise.ratio=0.1),
          generateNoise(x, "white", noise.ratio=0.3),
          generateNoise(x, "white", noise.ratio=0.5))
  }() |>
  {\(x) c(generateTimeShift(x, amount=1),
          generateTimeShift(x, amount=2))
  }()

# Create list of data and augmentations
data <- c(w, augmented)
```

The same approach can be used with a list of initial `Wave`-like objects.

```{r, echo=FALSE}
# Load some test audio data
f <- system.file("extdata", "AUDIOMOTH.flac", package="sonicscrewdriver")
w <- list(
  readAudio(f),
  readAudio(f),
  readAudio(f)
)

# Add noise to the audio data and then rotate the audio data using pipes and anonymous functions
augmented <- 
  w |>
  {\(x) c(generateNoise(x, "white", noise.ratio=0.1),
          generateNoise(x, "white", noise.ratio=0.3),
          generateNoise(x, "white", noise.ratio=0.5))
  }() |>
  {\(x) c(generateTimeShift(x, amount=1),
          generateTimeShift(x, amount=2))
  }()

# Create list of data and augmentations
data <- c(w, augmented)
```
