---
title: "Interacting with other tools"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, results=FALSE}
library(sonicscrewdriver)
```

# Audacity

Label files exported from [Audacity](https://www.audacityteam.org/) can be read into R using the `readAudacityLabels()` function. This function takes a path to a label file and returns a list of `Annotation` objects or a data frame with the start and end times of each label and the label text.

# AudioBlast

The [audioBlast API](https://api.audioblast.org) can be accessed using the `audioBlast()` function. Files can be downloaded from the API using the `audioBlastDownload()` function.

# AudioMoth

[AudioMoth](https://www.openacousticdevices.info/audiomoth) configuration files and wave files can be read into R using the `audiomothConfig()` and `audiomothWave()` functions, respectively. The `audiomothConfig()` function takes a path to a configuration file and returns a data frame of the configuration settings. The `audiomothWave()` function takes a path to a wave file and returns a list of extracted parameters.

# BirdNet
