---
title: "Interacting with other tools"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE}
library(sonicscrewdriver)
```

## Introduction

The `sonicscrewdriver` package provides a number of functions for interacting with other audio tools and software. This vignette provides an overview of the functions available for interacting with other tools.

## Interfaces with other tools

### Audacity

Label files exported from [Audacity](https://www.audacityteam.org/) can be read into R using the `readAudacityLabels()` function. This function takes a path to a label file and returns a list of `Annotation` objects or a data frame with the start and end times of each label and the label text. The `writeAudacityLabels()` function can be used to write a list of `Annotation` objects to an Audacity label file.

### AudioBlast

The [audioBlast API](https://api.audioblast.org) can be accessed using the `audioBlast()` function. Files can be downloaded from the API using the `audioBlastDownload()` function.

### AudioMoth

[AudioMoth](https://www.openacousticdevices.info/audiomoth) configuration files and wave files can be read into R using the `audiomothConfig()` and `audiomothWave()` functions, respectively. The `audiomothConfig()` function takes a path to a configuration file and returns a data frame of the configuration settings. The `audiomothWave()` function takes a path to a wave file and returns a list of extracted parameters.

### BirdNET Analyzer

[BirdNET Analyzer](https://birdnet.cornell.edu) is a deep learning model for the automatic detection of bird sounds. The `birdNetInstall()` function can be used to install the `ssd_birdnet` environment required to use the `birdNetAnalyse()` function with `SonicScrewdriver`. The `birdNetAnalyse()` function takes a list of sound files and analyses them using the `BirdNET-Analyzer`. The function either returns a data frame with the results of the analysis or a list of `Annotation` objects.

```{r, eval=FALSE}
# Install the Python environment
pythonInstall()
```

```{r, eval=FALSE}
# Analyse sound files using BirdNET-Analyzer
f <- system.file("extdata", "AUDIOMOTH.WAV", package="sonicscrewdriver")
birdNetAnalyse(f, lat=51.5, lon=0.1, date=as.Date("2021-01-01"))
```


### scikit-maad

[scikit-maad](https://scikit-maad.github.io/index.html) is an open source Python package dedicated to the quantitative analysis of environmental audio recordings. `sonicscrewdriver` makes use of the `reticulate` package to create a Python environment and to interact with scikit-maad. In `sonicscrewdriver` scikit-maad functions are prefixed with `maad_`.

```{r, eval=FALSE}
# Install the Python environment
pythonInstall()
```

```{r, eval=FALSE}
f <- system.file("extdata", "AUDIOMOTH.WAV", package="sonicscrewdriver")
w <- readWave(f)

maad_aci <- maad_acoustic_complexity_index(w)
```


## Exemplar applications

### Audio file with annotations from audioBlast to Audacity

The audio file of interest (id: [10754](https://bio.acousti.ca/node/10754)) is a recording of the Mole Cricket *Gryllotalpa vineae* from the [BioAcoustica](https://bio.acousti.ca) repository. Recordings and annotations from BioAcoustica are made available through the [audioBlast](https://audioblast.org) API.

First we will download the audio file and annotations from audioBlast.

```{r, results=FALSE}
# Find the file on audioblast
data <- audioblast("data", "recordings", source="bio.acousti.ca", id=10754)

# Download the file to the mole_cricket directory
audioblastDownload(data, dir="mole_cricket/")
```

This has downloaded the audio file to the `mole_cricket` directory. It has also saved a `metadata.csv` file that includes more information about the audio file retrieved from audioBlast.

Next, we will download the annotations for this file from audioBlast.

```{r, results=FALSE}
# Fetch annotations from audioBlast as `Annotation` objects
a <- audioblast("data", "annomate", source="bio.acousti.ca", id=10754, output="Annotations")
```

This has downloaded the annotations from audioBlast and converted them to a list of `Annotation` objects. Once annotation data are converted to `Annotation` objects they can be easily manipulated into a number of other useful formats.

In this case, we will convert the `Annotation` objects to an Audacity label file.

```{r}
# Convert the annotations to an Audacity label file
writeAudacityLabels(a, "mole_cricket/annotations.txt")
```

When we are finished, we can tidy our workspace by removing the `mole_cricket` directory.

```{r}
unlink("mole_cricket", recursive=TRUE)
```

## BirdNET Analyzer to Audacity label file

The `birdNetAnalyse()` function can be used to analyse sound files using the `BirdNET-Analyzer`.

Get the output of the BirdNET Analyzer as a list of `Annotation` objects.

```{r, eval=FALSE}
# Analyse sound files using BirdNET-Analyzer
f <- system.file("extdata", "AUDIOMOTH.WAV", package="sonicscrewdriver")
annotations <- birdNetAnalyse(f, output="Annotation")
```

Convert the `Annotation` objects to an Audacity label file.

```{r eval=FALSE}
# Convert the annotations to an Audacity label file
writeAudacityLabels(annotations, "birdnet_annotations.txt")
```

Cleaning up the workspace.

```{r, eval=FALSE}
unlink("birdnet_annotations.txt")
```
